---
title: "Gaia DR3 High Velocity"
author: "s"
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times

# The package "ezids" (EZ Intro to Data Science) includes a lot of the helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
library(ezids)
library(ggplot2)
library(plotly)
library(htmlwidgets)
# library(patchwork)
setwd("./GAIA-DR3-DATS6101/") 
```

```{r}
# read in v_tan_500 km_s-result.csv
df_high <- read.csv("v_tan_500km_s-result.csv")
df_high$sample <- "high"
df_general <- read.csv("general_sample_result.csv")
df_general$sample <- "general"
df <- rbind(df_high, df_general)

# calculate tangential velocity r_med_geo*sin(pm/1000*PI()/648000) * 978462
df$v_tan <- df$r_med_geo * sin(df$pm/1000*pi/648000) * 978462

# calculate absolute magnitudes from r_med_geo distance
df$Gmag_abs <- df$phot_g_mean_mag - 5 * log10(df$r_med_geo) + 5
df$RPmag_abs <- df$phot_rp_mean_mag - 5 * log10(df$r_med_geo) + 5
df$BPmag_abs <- df$phot_bp_mean_mag - 5 * log10(df$r_med_geo) + 5
```

```{r}
# columns of data
colnames(df)
```


```

```{r}
# make Gaia CMD of high velocity and general sample stars using phot_g_mean_mag and bp_rp columns using plotly. overlay the two CMDs on the same plot.
cmd <- ggplot(df, aes(x=bp_rp, y=Gmag_abs)) + 
  geom_point(aes(color=sample), alpha=0.25) + 
  scale_color_manual(values=c("red", "blue")) + 
  # invert y axis
  scale_y_reverse() +
  theme_bw() + 
  labs(x="BP-RP", y="G")

# ggplotly(cmd)

# save plotly figure
p <- ggplotly(cmd)
htmlwidgets::saveWidget(p, "cmd.html")
```

```{r}
# get number of rows in df with nan r_med_geo values
sum(!is.na(df$r_med_geo))

```

```{r}
```

```{r}


```


```{r}


```

```{r}
```

```{r}
```

